---
title: "Gender balance between major and minor characters"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r echo=F,eval=F}
try(setwd("~/OneDrive - Cardiff University/Research/Cardiff/VideoGameScripts/project/analysis/"))
```


# Introduction

This report explores whether the gender bias in the data as a whole is also observed when looking at either only major characters or only minor characters. Major characters may be playable characters, party characters, characters more central to the plot, or characters that speak more. One concern is that dialogue data for major characters is more complete and that gender is easier to code. This might bias estimates of the proportion of dialogue by female characters. 

The analyses below demonstrate that conclusions about the proportion of female dialogue are unlikely to be affected by such concerns in our data.

# Gender bias in the dialogue of major and minor characters

We first look at major and minor characters as discrete groups.

Load libraries:

```{r warning=F, message=F}
library(rjson)
library(ggplot2)
library(ggstance)
library(mgcv)
library(knitr)
library(betareg)
```

Load data (for games with coded main player characters):

```{r}
folders = list.dirs("../data", recursive = T)
folders = folders[sapply(folders,function(X){
    "stats_by_character.csv" %in% list.files(X)
  })]


allGames = NULL
for(folder in folders){
  shortName = tail(strsplit(folder,"/")[[1]],1)
  js = fromJSON(file = paste0(folder,"/meta.json"))
  alternativeMeasure = FALSE
  if(!is.null(js$alternativeMeasure)){
    alternativeMeasure = js$alternativeMeasure
  }
  if(!alternativeMeasure){
    statsByChar = read.csv(paste0(folder,"/stats_by_character.csv"),stringsAsFactors = F)
    statsByChar = statsByChar[!is.na(statsByChar$words),]
    statsByChar = statsByChar[statsByChar$words>0,]

    if(nrow(statsByChar)>0 && !is.null(js$mainPlayerCharacters)){
      majc = statsByChar$charName %in% js$mainPlayerCharacters
      minc = (!statsByChar$charName %in% js$mainPlayerCharacters) & 
        (statsByChar$group %in% c("male","female"))
      majc.Female = statsByChar$group=="female" & majc
      majc.Male = statsByChar$group=="male" & majc
      minc.Female =  statsByChar$group=="female" & minc
      minc.Male =  statsByChar$group=="male" & minc
      
      # Only include games with coded major characters
      if((sum(majc.Female) + sum(majc.Male))> 0 ){
        print(folder)
        majc.Female.words = sum(statsByChar[majc.Female,]$words)
        propFemaleDialogue.mainChar = 0
        if(majc.Female.words>0){
         propFemaleDialogue.mainChar = 
           sum(statsByChar[majc.Female,]$words) / 
             (sum(statsByChar[majc.Female,]$words) + 
              sum(statsByChar[majc.Male,]$words)) 
        }
        propFemaleDialogue.minorChar = 
          sum(statsByChar[minc.Female,]$words) /
            (sum(statsByChar[minc.Female,]$words) +
             sum(statsByChar[minc.Male,]$words))
        
        ret = data.frame(
          folder = folder,
          game = js$game,
          shortName = shortName,
          group = c("major","minor"),
          numFemaleWords = c(
            sum(statsByChar[majc.Female,]$words),
            sum(statsByChar[minc.Female,]$words)),
          numMaleWords = c(
            sum(statsByChar[majc.Male,]$words),
            sum(statsByChar[minc.Male,]$words)),
          propFemaleDialogue = c(
            propFemaleDialogue.mainChar,
            propFemaleDialogue.minorChar
          ),
          numFemaleCharacters = c(sum(majc.Female),sum(minc.Female)),
          numMaleCharacters = c(sum(majc.Male),sum(minc.Male))
        )
        allGames = rbind(allGames, ret)
      }
    }    
  }
}
```

Visualise total amount of female vs. male dialogue for major and minor characters:

```{r}

allGames.Maj.PercentFemale =
  100 * (
  sum(allGames[allGames$group=="major",]$numFemaleWords) /
   (sum(allGames[allGames$group=="major",]$numFemaleWords) +
      sum(allGames[allGames$group=="major",]$numMaleWords)))
allGames.Maj.PercentMale = 100 - allGames.Maj.PercentFemale

allGames.Min.PercentFemale =
  100 * (
  sum(allGames[allGames$group=="minor",]$numFemaleWords) /
   (sum(allGames[allGames$group=="minor",]$numFemaleWords) +
      sum(allGames[allGames$group=="minor",]$numMaleWords)))
allGames.Min.PercentMale = 100 - allGames.Min.PercentFemale

allGames$propFemaleCharacters= 
  allGames$numFemaleCharacters / 
    (allGames$numFemaleCharacters + allGames$numMaleCharacters)

dx = data.frame(
  Gender=factor(c("Male","Female","Male","Female"),
                levels=c("Male","Female")),
  Group = factor(c("Major","Major","Minor","Minor"),
                 levels=c("Minor","Major")),
  percentageWords=
    c(allGames.Maj.PercentMale,allGames.Maj.PercentFemale,
      allGames.Min.PercentMale,allGames.Min.PercentFemale))

ggplot(dx,aes(x=Group,y=percentageWords,fill=Gender))+ geom_bar(stat='identity')+
  geom_hline(yintercept=50,linetype="dotted") +
  coord_flip(ylim = c(0,100)) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "top") +
  scale_fill_discrete(breaks=c("Female","Male"),name="Gender")+
  ylab("% Words Spoken") +
  xlab("")
```

Below is a boxplot of the distribution of individual games. It's clear that the minor character group has a higher mean proportion of female dialogue, though the range for major characters is higher.

```{r}
ggplot(allGames, aes(y=propFemaleDialogue,x=group,colour=group)) + 
  geom_boxplot() + 
  geom_jitter() +
  theme(legend.position = "none")+
  xlab("Character type") +
  ylab("Proportion of female dialogue")
```

Statistical test of average proportion of female dialogue in each game, comparing major and minor characters.

```{r}
t.test(allGames$propFemaleDialogue~allGames$group)
```

The test is significant, suggesting that the female dialogue is lower in major characters than in minor characters.

However, we also know that the proportion of female dialogue is predicted by the proportion of female characters. The regression below tests whether the character groups (major or minor) predict the proportion of female dialogue over and above the proportion of female characters within the group:

```{r}
mInt = lm(propFemaleDialogue~ group * propFemaleCharacters, data = allGames)
summary(mInt)
```

The effect of the proprotion of female characters is significant, but the effect of group is not, nor is the interaction. 

```{r}
ggplot(allGames, 
       aes(x=propFemaleCharacters,
           y=propFemaleDialogue,
           colour=group)) +
  geom_point() +
  stat_smooth(method='lm') +
  xlab("Proportion of female characters in group") +
  ylab("Proportion of female dialogue in group")
```

In summary, the gender bias against female dialogue is evident for both major and minor characters. This bias is exaggerated for major characters. However, this can be mostly explained by the number of female characters, rather than a systematic difference between major and minor character groups.

\clearpage
\newpage

The full list of results:

```{r}
kable(allGames[,c("game","group","propFemaleDialogue")])
```


\clearpage
\newpage


## Bias in different quantiles of character dialgoue

Does the gender bias differ for characters that speak a lot compared to characters that don't?

For each game, we divide characters into four groups based on the amount of dialogue they speak. Characters are ranked by the proportion of dialogue that they speak within the game. Then the characters are split into four groups of even number (the 'quantiles' of the dialogue proportions). Across all games, the total number of words is calculated for each gender for each quantile. 

```{r}
folders = list.dirs("../data", recursive = T)
folders = folders[sapply(folders,function(X){
    "stats_by_character.csv" %in% list.files(X)
  })]


allChars = NULL
for(folder in folders){
  shortName = tail(strsplit(folder,"/")[[1]],1)
  js = fromJSON(file = paste0(folder,"/meta.json"))
  alternativeMeasure = FALSE
  if(!is.null(js$alternativeMeasure)){
    alternativeMeasure = js$alternativeMeasure
  }
  if(!alternativeMeasure){
    statsByChar = read.csv(paste0(folder,"/stats_by_character.csv"),stringsAsFactors = F)
    statsByChar = statsByChar[!is.na(statsByChar$words),]
    statsByChar = statsByChar[statsByChar$words>0,]
    if(nrow(statsByChar)>0){
        statsByChar = statsByChar[statsByChar$group %in% c("male","female"),]
        statsByChar$dialogProp = statsByChar$words/sum(statsByChar$words)
        allChars = rbind(allChars,statsByChar)
      }
    }    
}
```


```{r}
numQuantiles = 4
allChars$Quantile = cut(allChars$dialogProp,
      breaks= quantile(allChars$dialogProp,
          probs = seq(0,1,length.out=numQuantiles+1)))
q = data.frame(
  Quantile = 1:numQuantiles,
  femaleWords = tapply(
    allChars[allChars$group=="female",]$words,
    allChars[allChars$group=="female",]$Quantile,sum),
  maleWords = tapply(
    allChars[allChars$group=="male",]$words,
    allChars[allChars$group=="male",]$Quantile,sum),
  femaleChars = tapply(
    allChars[allChars$group=="female",]$words,
    allChars[allChars$group=="female",]$Quantile,length),
  maleChars = tapply(
    allChars[allChars$group=="male",]$words,
    allChars[allChars$group=="male",]$Quantile,length)
)
q$propFemale = q$femaleWords/ (q$femaleWords+q$maleWords)
q$propMale = q$maleWords/ (q$femaleWords+q$maleWords)
q$propFemaleChar = q$femaleChar/ (q$femaleChar+q$maleChar)
q$propMaleChar = q$maleChar/ (q$femaleChar+q$maleChar)
```

There are small but significant differences in the proportion of words in each of the four quantiles:

```{r}
q[,c("femaleWords","maleWords","propFemale","propMale")]
chisq = chisq.test(q[,c("femaleWords","maleWords")])
chisq
pv = chisq$p.value
if(pv < 0.0001){
  pv = paste0("p = 0.0001")
  } else{
  pv = paste0("p = ",round(pv,3))
}
chisqOut = paste0("$\\chi^2$ = ", round(chisq$statistic,2),", ",pv)
cat(chisqOut,file="../results/latexStats/quantileWordsChiSq.tex")
# compare just first and last quantile
chisq.test(q[c(1,4),c("femaleWords","maleWords")])
```

However, the proportion of characters is not significantly different (the p-value is marginal at the 0.05 level, but very different from the result above).

```{r}
q[,c("femaleChars","maleChars","propFemaleChar","propMaleChar")]
chisq = chisq.test(q[,c("femaleChars","maleChars")])
pv = chisq$p.value
if(pv < 0.0001){
  pv = paste0("p = 0.0001")
  } else{
  pv = paste0("p = ",round(pv,3))
}
chisqOut = paste0("$\\chi^2$ = ", round(chisq$statistic,2),", ",pv)
cat(chisqOut,file="../results/latexStats/quantileCharChiSq.tex")
# compare just first and last quantile
chisq.test(q[c(1,4),c("femaleChars","maleChars")])
```

Plot results:

```{r}
q2 = data.frame(
  Quantile = paste0("Q",rep(1:numQuantiles,2)),
  gender = factor(rep(c("female","male"),each=numQuantiles),
                  levels=c("male","female")),
  prop = c(q$propFemale,q$propMale),
  propChar =c(q$propFemaleChar,q$propMaleChar)
)

ggplot(q2,aes(x=Quantile,fill=gender,y=prop)) +
  geom_bar(stat = "identity",position="stack") +
  ylab("Proportion of dialogue") +
  xlab("Quantile (fewest words to most words)") +
  ggtitle("Proportion of words spoken by quantile")

ggplot(q2,aes(x=Quantile,fill=gender,y=propChar)) +
  geom_bar(stat = "identity",position="stack") +
  ylab("Proportion of characters") +
  xlab("Quantile (fewest words to most words)")+
  ggtitle("Proportion of characters by quantile")
  

```

In summary, there are small differences in the proportion of female dialogue for talkative and less talkative characters. But the overall 


\clearpage
\newpage

# Other estimates of bias

One concern is that the coding for main characters may be more complete or more accurate than coding for minor characters, because minor characters are harder to find in videos, have less documentation on wikis and less direct linguistic cues. 

## Method 1: Complete vs. incomplete scripts

Each game is coded for whether it's complete, partial or a sample of the dialogue in the full game. If the estimate was biased by the completeness of the coding, then we might expect to see a difference in the estimated proportion of female dialogue between these types.

Below is a boxplot showing how the proportion of female dialogue differs between the game types.

First, we load the data:

```{r}
allGames.completeness = NULL
for(folder in folders){
  js = fromJSON(file = paste0(folder,"/meta.json"))
  completeness = NA
  if(!is.null(js$sourceFeatures)){
    if(!is.null(js$sourceFeatures$completeness)){
      completeness = js$sourceFeatures$completeness
    }
  }
  alternativeMeasure = FALSE
  if(!is.null(js$alternativeMeasure)){
    alternativeMeasure = js$alternativeMeasure
  }
  if(!alternativeMeasure){
    stats = read.csv(paste0(folder,"/stats_by_character.csv"),stringsAsFactors = F)
    propFemaleDialogue = sum(stats[stats$group=="female",]$words) /
      sum(stats[stats$group %in% c("male","female"),]$words)
    propFemaleCharacters = sum(stats$group=="female")/
      sum(stats$group %in% c("male","female"))
    ret = data.frame(
      folder = folder,
      game = js$game,
      completeness = completeness,
      propFemaleDialogue = propFemaleDialogue,
      propFemaleCharacters = propFemaleCharacters,
      year = js$year
    )
    allGames.completeness = rbind(allGames.completeness, ret)
  }
}
allGames.completeness = allGames.completeness[
  is.finite(allGames.completeness$propFemaleDialogue),]
```

Below we plot the proportion of female dialogue according to the extent of the data coded:

```{r}
ggplot(allGames.completeness[!is.na(allGames.completeness$completeness),],
       aes(x=completeness, y =propFemaleDialogue, colour=completeness)) +
  geom_boxplot() +
  geom_jitter() +
  theme(legend.position = "none")+
  xlab("Extent of the dialogue data in the corpus") +
  ylab("Proportion of female dialogue") +
  coord_cartesian(ylim=c(0,1))
```

A t-test comparing proportion of female dialogue in 'complete' and 'high' sources:

```{r}
t.test(propFemaleDialogue~completeness,
       data = allGames.completeness[
         allGames.completeness$completeness %in% 
           c("complete", "high"),])
```


The proportion of female dialogue is slightly higher for more complete data. This might suggest that the estimate of female dialogue is underestimated in the corpus. However, we also know that the proportion of female dialogue is predicted by the proportion of female characters:

```{r}
ggplot(allGames.completeness[!is.na(allGames.completeness$completeness),],
       aes(x=propFemaleCharacters, 
           y =propFemaleDialogue,
           colour=completeness)) +
  geom_point() +
  stat_smooth(method='lm') +
  xlab("Proportion of female characters in group") +
  ylab("Proportion of female dialogue in group")
```

It could also be the case that the year of publication is related to how complete the script is (since earlier games have less content, and are more likely to have accessible data). This appears to be the case:

```{r}
ggplot(allGames.completeness[!is.na(allGames.completeness$completeness),],
       aes(x=year, y=completeness,colour=completeness)) +
  geom_boxplot() +
  geom_jitter()
```


This makes it seem like the the relationship with completeness may be partially driven by the proprotion of female characters and/or by the date of release. Here is a regression, predicting the proportion of female dialogue by completeness and by the proportion of female characters:

```{r}
mComp = lm(propFemaleDialogue ~ completeness + propFemaleCharacters + year,
   data = allGames.completeness)
summary(mComp)
```

It seems there is no effect of completeness once the proprotion of female charaters is taken into account.

\clearpage
\newpage

## Method 2: Cumulative gender balance

We can try to estimate how our measuring of gender balance is affected by the coding of minor characters. This can be done by looking at how the estimate varies as we observe more and more minor characters. That is, what would the estimate be like if we had only coded the top 10% of most prolific characters (those that speak most), or the top 20%, or top 30% etc. At some point, the estimate of gender balance would converge on the final estimate for all coded characters.


### Example

Here's a function that works out the gender balance in dialogue over the range of characters. It ranks all characters from most words to least words, then works out the gender balance taking into account just the top character, the top two characters, the top three characters etc.

```{r}
getBalanceOverCumulativeCharacterRange = function(statsByChar){
  # Remove other groups except male and female
  statsByChar = statsByChar[statsByChar$group %in% c("male","female"),]
  statsByChar = statsByChar[!is.na(statsByChar$words),]
  statsByChar = statsByChar[statsByChar$words>0,]

  # Sorted list of number of words for each character (most to least)
  sortedUniqueNumOfWords = sort(unique(statsByChar$words),decreasing = TRUE)
  
  # Table of total number of words observed for each gender 
  #  as the number of characters observed increases from
  #  character with most dialogue to least dialogue
  wordsByGender.Cumulative = 
    sapply(sortedUniqueNumOfWords,
      function(minNumWords){
        x = statsByChar[statsByChar$words>=minNumWords,]
        femaleWords = sum(x[x$group=="female",]$words)
        maleWords = sum(x[x$group=="male",]$words)
        femaleProp = femaleWords / (femaleWords+maleWords)
        return(c(femaleWords,maleWords))
      })
  # Convert to proportion of female dialogue
  femalePropCumulative = wordsByGender.Cumulative[1,] /
    colSums(wordsByGender.Cumulative)
  totalEstimate = femalePropCumulative[length(femalePropCumulative)]
  # Binomial test at each point: is it significantly different
  #  from the total?
  sigDifferentFromTotal.p = apply(wordsByGender.Cumulative,2,
    function(wbg){
      x = binom.test(wbg,p = totalEstimate)$p.value
  })
  
  sigDifferentFromTotal = sigDifferentFromTotal.p>0.05
  firstNonSignif = which(sigDifferentFromTotal)[1]
  # stable non-signif (point after last significant result)
  firstStableNonSignif = length(sigDifferentFromTotal) - which(!rev(sigDifferentFromTotal))[1] + 2
  
  numCharCumulative = sapply(sortedUniqueNumOfWords,
      function(minNumWords){
        sum(statsByChar$words>=minNumWords)
      })
  
  numCharBeforeEstimateFirstNotSigDiff = numCharCumulative[firstNonSignif]
  numCharBeforeEstimateStabilises = numCharCumulative[firstStableNonSignif]
  
  return(list(femalePropCumulative = femalePropCumulative,
      numCharCumulative = numCharCumulative,
      p = sigDifferentFromTotal.p,
      numCharBeforeEstimateFirstNotSigDiff = numCharBeforeEstimateFirstNotSigDiff,
      numCharBeforeEstimateStabilises = numCharBeforeEstimateStabilises
              ))
}

# Function for visualising the information
plotCumulativeCharRange = function(stats){
  plot(stats$femalePropCumulative~
       stats$numCharCumulative,
     xlab = "Number of characters observed",
     ylab = "Proportion of female dialogue",
     ylim=c(0,1),
     col = 1 + (stats$p>0.05))
abline(h=tail(stats$femalePropCumulative,n=1), 
       col=rgb(1,0,0,0.5))
points(stats$numCharBeforeEstimateStabilises,
       tail(stats$femalePropCumulative,n=1) + 0.1,
       pch=5, col="red")
}

```

We can apply this to Chrono Trigger:

```{r}
chrono =  read.csv("../data/ChronoTrigger/ChronoTrigger/stats_by_character.csv",
                   stringsAsFactors = F)
chrono = chrono[!is.na(chrono$words),]
chrono = chrono[chrono$words>0,]
chrono.stats = getBalanceOverCumulativeCharacterRange(chrono)
plotCumulativeCharRange(chrono.stats)
```

The points in the plot above shows the gender balance when taking into account various numbers of characters. Note that there is not an estiamte for every possible number of characters, since several characters are tied in the number of words they speak. The red line shows the gender balance in the full game. The points are coloured red if they are not significantly different from the gender balance in the full game.

It's clear that the proportion of female dialogue is higher for characters with more dialogue. The first time that the estimate is not signficantly different from the total estimate is after seeing `r chrono.stats$numCharBeforeEstimateFirstNotSigDiff` characters, or after seeing `r round(100* chrono.stats$numCharBeforeEstimateFirstNotSigDiff / tail(chrono.stats$numCharCumulative,n=1))`% of the characters. After this, the estimate does not change significantly. This is indicated with a red diamond on the plot. This also happens to be the same as the point at which the estimate stabilises (the point after which the estimate is never again significantly different from the total estimate).

Things are a little different for Final Fantasy XII, which has many more characters:

```{r}
FFXII =  read.csv("../data/FinalFantasy/FFXII_B/stats_by_character.csv",
                  stringsAsFactors = F)
FFXII.stats = getBalanceOverCumulativeCharacterRange(FFXII)
plotCumulativeCharRange(FFXII.stats)
```

Here, it's clear that the proportion of female dialogue is *lower* for characters with more dialogue. The first time that the estimate is not signficantly different from the total estimate is after seeing `r FFXII.stats$numCharBeforeEstimateFirstNotSigDiff` characters, or after seeing `r round(100* FFXII.stats$numCharBeforeEstimateFirstNotSigDiff / tail(FFXII.stats$numCharCumulative,n=1))`% of the characters.

Things are different again for Daggerfall:

```{r}
Daggerfall =  read.csv("../data/ElderScrolls/Daggerfall/stats_by_character.csv",
                  stringsAsFactors = F)
Daggerfall.stats = getBalanceOverCumulativeCharacterRange(Daggerfall)
plotCumulativeCharRange(Daggerfall.stats)
```

Here, the female dialogue is over-estimated for characters with lots of dialogue, but then under-estimated for mid-range characters. It is only after seeing `r round(100* Daggerfall.stats$numCharBeforeEstimateFirstNotSigDiff / tail(Daggerfall.stats$numCharCumulative,n=1))`% of characters that the estimate stabilises.

\clearpage
\newpage

## Cumulative gender balance over all games

We can now estiamte the statistics above for all games:

```{r}
folders = list.dirs("../data", recursive = T)
folders = folders[sapply(folders,function(X){
  "stats_by_character.csv" %in% list.files(X)
})]


allGames.Cum = NULL
for(folder in folders){
  shortName = tail(strsplit(folder,"/")[[1]],1)
  js = fromJSON(file = paste0(folder,"/meta.json"))
  alternativeMeasure = FALSE
  if(!is.null(js$alternativeMeasure)){
    alternativeMeasure = js$alternativeMeasure
  }
  if(!alternativeMeasure){
    print(folder)
    statsByChar = read.csv(paste0(folder,"/stats_by_character.csv"),stringsAsFactors = F)
    statsByChar = statsByChar[!is.na(statsByChar$words),]
    statsByChar = statsByChar[statsByChar$words>0,]

    if(nrow(statsByChar)>0){
      gameStats = getBalanceOverCumulativeCharacterRange(statsByChar)
      totalNumChar = tail(gameStats$numCharCumulative,n=1)
      percentCharBeforeEstimateStabilises= 
        100 * (gameStats$numCharBeforeEstimateStabilises / totalNumChar)
      ret = data.frame(
        folder = folder,
        game = js$game,
        shortName = shortName,
        totalNumChar = totalNumChar,
        femalePercentCumulative = 100*gameStats$femalePropCumulative,
        percentCharCumulative = 100*(gameStats$numCharCumulative/totalNumChar),
        percentCharBeforeEstimateStabilises = percentCharBeforeEstimateStabilises
      )
      ret$femalePercentTotal = tail(ret$femalePercentCumulative,n=1)
      allGames.Cum = rbind(allGames.Cum,ret)
    }    
  }
}
allGames.Cum$diffFromTotalEstimate = 
  allGames.Cum$femalePercentCumulative - allGames.Cum$femalePercentTotal
```


The graph below visualises the data for all games. The vertical axis plots the distance from the final total estiamte for the specific game, so that all games converge on zero (no difference from the total estimate). The horizontal axis shows the percentage of characters seen, so that each game is normalised for the number of characters. Red diamonds show the points at which the estimate for each game stabilises. The boxplot shows the distribution of the stabilisation points over the horizontal axis.  

```{r}
ggplot(allGames.Cum,
       aes(x=percentCharCumulative,
           y=diffFromTotalEstimate,
           group=folder)) +
  geom_line(alpha=0.3) + 
geom_hline(yintercept = 0, alpha=0.5,colour='red') +
  geom_point(data=allGames.Cum[!duplicated(allGames.Cum$folder),], 
    aes(x=percentCharBeforeEstimateStabilises,
        y=15),
    colour="red",shape =5) +
  geom_boxploth(data=allGames.Cum[!duplicated(allGames.Cum$folder),],
               aes(x=percentCharBeforeEstimateStabilises,
                   y=25,group=NULL), colour="red", width=8) +
  ylab("Difference from total female proportion of dialogue") + 
  xlab("% of characters observed")
```

```{r echo=F}
qq = quantile(allGames.Cum[!duplicated(allGames.Cum$folder),]$percentCharBeforeEstimateStabilises,probs = c(0.025,0.975))
```


Estimates for all games stabilised before seeing 95% of their characters, suggesting that the estimates for the specific games in the corpus would not change much with the addition of data from more minor characters.

The mean stabilisation point is after seeing `r mean(allGames.Cum[!duplicated(allGames.Cum$folder),]$percentCharBeforeEstimateStabilises)`% of characters (95% quantile = [`r qq[1]` ,`r qq[2]`].


## Is the estimate biased?

The graphs above show that, when considering only the most prolific characters, there are games which both over- and under- estimate the proprotion of female dialogue.  Additionally, we want to formally test whether the estimate of the estimate of the proportion of female dialogue is biased *in the corpus as a whole*. We can do this by looking at the data after seeing 10% of the characters in a game. The t-test below tests whether there is a bias one way or the other:

```{r}
firstTenPercent = allGames.Cum[allGames.Cum$percentCharCumulative>=10.00,]
firstTenPercent = firstTenPercent[!duplicated(firstTenPercent$folder),]
tBias = t.test(firstTenPercent$diffFromTotalEstimate)
tBias
```

The mean difference from the final estimate is `r tBias$estimate`, which indicates that the estimate based on prolific characters may be under-estimating the proportion of female speech on average. However, this is not significantly different from 0.

A similar point is made when drawing boxplots for the mean estimate of female dialogue % for each 10% of the range of characters seen (the whiskers show the full range of the data). Seeing only 10% of the characters biases the estimates (the proportion of female dialogue is under-estimated, though this isn't significant as shown by the t-test above). After seeing 10% of the characters, the estimates are very close to the final estimates after seeing all the characters.

```{r}
allGames.Cum$cut = cut(allGames.Cum$percentCharCumulative,
                        breaks = seq(0,100,by=10))
ggplot(allGames.Cum[!duplicated(paste(allGames.Cum$folder, allGames.Cum$cut)),],
       aes(y=diffFromTotalEstimate,
           x = cut)) +
  geom_boxplot(alpha=0.5, coef = Inf) +
  stat_summary(fun=mean, geom="point", shape=20, size=2, color="red", fill="red")
```

However, cutting the data in 10% chunks is arbitrary, and the data within each chunk are not necessarily normally distributed.

A more generalised answer can be give by fitting a general additive model (GAM), predicting the difference from the total estimate based on the percentage of characters seen, with random effects to capture the dependency of datapoints that belong to the same game. 

```{r}
mBeta = betareg(I(0.3 + (diffFromTotalEstimate/100)) ~
                percentCharCumulative | percentCharCumulative,
                data = allGames.Cum[allGames.Cum$percentCharCumulative>5,])

summary(mBeta)

px = predict(mBeta, newdata = data.frame(
  percentCharCumulative = seq(0,100,by=5)
))
px.var = predict(mBeta, type="variance", newdata = data.frame(
  percentCharCumulative = seq(0,100,by=5)
))

allGames.Cum$folder.factor = factor(allGames.Cum$folder)
mGAM = gam(diffFromTotalEstimate~
             s(percentCharCumulative) +
             s(folder.factor, bs = 're'), 
           data = allGames.Cum[allGames.Cum$percentCharCumulative>5,])
summary(mGAM)
```

The model suggests that there is a significant relationship overall. We can visualise the curve to get an idea of the trends:

```{r}
plot.gam(mGAM, xlab="Percentage of characters seen",
     ylab = "Difference from final % of female dialogue",
     select=1)
abline(h=0,col=1)
abline(v=seq(0,100,by=10), lty=2,col="gray")
```

In contrast to the boxplot, the visualisation suggets that female dialogue tends to be *over-estimated* for prolific characters, then slightly under-estimated after seeing about between 20-50% of characters. After about 60% of characters, the estimate stabilises. Note that the range is very small - predicting biases of around 1% at most.


The plot below shows how the stabilisation point varies with the total number of characters in the game. There's no strong relationship between the two.

```{r}
ggplot(data=allGames.Cum[!duplicated(allGames.Cum$folder),],
       aes(y=percentCharBeforeEstimateStabilises,
           x =totalNumChar)) +
  geom_point()

cor.test(allGames.Cum[!duplicated(allGames.Cum$folder),]$percentCharBeforeEstimateStabilises,
         allGames.Cum[!duplicated(allGames.Cum$folder),]$totalNumChar)

```

\clearpage
\newpage

# Conclusion

The analyses above suggest that the gender bias in video game dialogue is present in major and minor characters. Although the bias was stronger for major characters, this can be mostly explained by the low number of female characters in this group, rather than a systematic difference between major and minor character groups *per-se*.

Furthermore, the estimate of the proportion of female characters is not systematically biased across games due to being able to get more complete or accurate data on more prolific characters. Individual games were biased in different directions, but with the majority converging after seeing 90% of the characters. The estimate of female dialogue across all games was not different from the final estimate after seeing the top 60% prolific characters in each game. For some ranges of the data, the average estimate of female dialogue across all games was an *under-estimate*, rather than an over-estimate. In any case, the bias in the estimate after seeing 20% of the characters are in the range of a few percentage points. This is much smaller than the overall gender bias in the corpus.

In conclusion, it's unlikely that the gender biases reported in the main data are affected by a tendency to have more complete or accurate data on more central characters.
